---
title: "Tut_2_Code"
author: "Ricky Fung"

format: html
---

#### Preamble

# Purpose: determines the number of reports based on the category of crime.
# Date: 2024/01/15
# Contact: ricky.fung\@mail.utoronto.ca

#### Workspace setup

# load packages

```{r}
#| message: false
#| warning: false

library(janitor)
library(tidyverse)
library(dplyr)
library(opendatatoronto)
```

# get package

```{r}
package <- show_package("police-annual-statistical-report-reported-crimes")
```

# get all resources for this package

```{r}
resources <- list_package_resources("police-annual-statistical-report-reported-crimes")
```

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources

```{r}
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
```

# load the first datastore resource as a sample

```{r}
data <- filter(datastore_resources, row_number()==1) %>% get_resource()
head(data)
```

## Cleaning and tidying data

# Clean column names

```{r}
cleaned <-
  clean_names(data)
```

# Select necessary columns

```{r}
cleaned <-
  cleaned |>
  select(
    id,
    category
  )

head(cleaned)
```

# Save cleaned data as csv file

```{r}
write_csv(x=cleaned, file="reported_crimes_reports.csv")
```

## Plot

# Create bar graph to find the number of reported crimes for each category in 2014

```{r}
cleaned |>
  ggplot(aes(x = category)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +  # Rotate x-axis labels
  labs(title = "Number of Reported Crimes in 2014 by Crime Category", x = "Categories", y = "Number of Reports")

```

## Citations

```{r}
citation()
citation("janitor")
citation("tidyverse")
citation("dplyr")
citation("opendatatoronto")
```
