---
title: "Tut_2_Code"
author: "Ricky Fung"

format: html
---

#### Preamble

# Purpose: To examine the number of reported crimes based on the category of crime in the year
# 2014 and create a graph.
# Date: 2024/01/15
# Contact: ricky.fung@mail.utoronto.ca

#### Workspace setup

# load packages

```{r}
#| message: false
#| warning: false

library(janitor)
library(tidyverse)
library(dplyr)
library(opendatatoronto)
```

# get package

```{r}
package <- show_package("police-annual-statistical-report-reported-crimes")
```

# get all resources for this package

```{r}
resources <- list_package_resources("police-annual-statistical-report-reported-crimes")
```

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources

```{r}
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
```

# load the first datastore resource as a sample

```{r}
data <- filter(datastore_resources, row_number()==1) %>% get_resource()
head(data)
```

## Cleaning and tidying data

# Clean column names

```{r}
cleaned <-
  clean_names(data)
```

# Select necessary columns

```{r}
cleaned <-
  cleaned |>
  select(
    id,
    category
  )

head(cleaned)
```

# Save cleaned data as csv file

```{r}
write_csv(x=cleaned, file="reported_crimes_reports.csv")
```

## Plot

# Create bar graph to find the number of reported crimes for each category in 2014

```{r}
cleaned |>
  ggplot(aes(x = category)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +  # Rotate x-axis labels
  labs(title = "Number of Reported Crimes in 2014 by Crime Category", x = "Categories", y = "Number of Reports")

```
```{r}
cleaned |>
  count(category)
```

# Share
The 6 categories of crime we are interested in are Controlled Drugs and Substances Act, Crimes Against Property, Crimes Against the Person, Criminal Code Traffic, Other Criminal Code Violations, Other Federal Statute Violations. We are interested in the number of reports made for each category.

We found that the crimes: Controlled Drugs and Substances Act had 1696 reports, Crimes Against Property had 15392, Crimes Against the Person had 8423, Criminal Code Traffic had 1832, Other Criminal Code Violations had 3737, and Other Federal Statute Violations had 920 reports.

## Citations

```{r}
citation()
citation("janitor")
citation("tidyverse")
citation("dplyr")
citation("opendatatoronto")
```
